# Biological Variation Report for (`r input$analyte`)


```{r hello-random, echo=FALSE}
pxtablek <- function(x, ...) { 
	if(class(x)=='list'){
  		print(xtable::xtable(x[[1]], digits=x$digits, caption=paste('</br>',x$caption)  ),type='html', ...) 
  	}else{
  		print(xtable::xtable(x),type='html', ...) 
  	}
}
opts_chunk$set(echo=FALSE, results='asis',fig.path='tempFiguresForKnitrReport/', fig.width=9, fig.height=7)
```


This report was created using the **BV** tool for analyzing biological variation data. **BV** is developed by **Turcosa Analytics** and can be accessed online at http://opensoft.turcosa.com.tr/BV 

Additional documentation for the **BV**, including instructions on how to download the application for offline use, can be found at https://github.com/selcukorkmaz/BV


# Introduction 

In this report, we analyzed the biological analyte by following the steps provided by Braga and Panteghini (2017). First, we applied 3 different outlier detection steps: (i) outliers in the sets of duplicate results, (ii) outliers in the variances of the results from each subject and  (iii) outliers in the variances of the results from each subject. We used Cochran test for the first two and Reed's criterion for the last one. Then we applied normality tests using Shapiro-Wilk and Kolmogorov-Smirnov tests in two steps: (i) on set of results from each individual, (ii) on mean values of subjects. Next, we applied a subset analysis to compare (i) means and (i) average within-subject total variances of gender groups. Finally, we performed analysis of variance and provided coefficient of variation results, ANOVA table and quality measures for all subjects, males and females separately.

Here are the abbreviations we used in this report:

**CV:** coefficient of variation
**CV_A:** analytical CV
**CV_I:** within CV
**CV_G:** between CV
**II:** index of individuality




*******
# 1. Outlier Detection

## 1.1. Step 1: Outliers in the sets of duplicate results using the Cochran test

```{r}
datatable(outlierResults()[[1]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
```
*******


## 1.2. Step 2: Outliers in the variances of the results from each subject using the Cochran test

```{r}
datatable(outlierResults()[[2]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
```
*******


## 1.3 Step 3: Outliers in the variances of the results from each subject using the Cochran test

```{r}
datatable(outlierResults()[[3]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
```
*******

# Normality Test
## 2.1. Step 1: On set of results from each individual
```{r}
      datatable(normalityResults()[[1]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))

```

## 2.2. Step 2: On mean values of subjects
```{r}
      datatable(normalityResults()[[2]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))

```

# 3. Subset Analysis
## 3.1. Step 1: Student's t test for mean differences of gender
```{r}
      datatable(subsetResults()[[2]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))

```

## 3.2. Step 2: Student's t test for average within-subject total variance
```{r}
      datatable(subsetResults()[[4]], extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))

```

# 4. Analysis of Variance
## 4.1. All subjects



### 4.1.1. Step 1: Coefficient of variation results

```{r}
 res = analysisOfVarinceResults()
    
    if(input$run && input$showResults == "original"){
      
      result = datatable(res$resultsAllOriginal$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
    
    
     if(input$run && input$showResults == "transformed"){
      
      result = datatable(res$resultsAllTransformed$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
    
    
     if(input$run && input$showResults == "transformBack"){
      
      result = datatable(res$resultsAllTransformBack$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
     }

result
```

### 4.1.2. Step 2: Analysis of variance table

```{r}
    
    res = analysisOfVarinceResults()
    
    if(input$run && input$showResults == "original"){
      
      result = datatable(res$resultsAllOriginal$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
    
    
     if(input$run && input$showResults == "transformed"){
      
      result =  datatable(res$resultsAllTransformed$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
    
    
     if(input$run && input$showResults == "transformBack"){
      
      result = datatable(res$resultsAllTransformBack$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
     }

result
    
    
```


### 4.1.3. Step 3: Quality measures

```{r}
 res = analysisOfVarinceResults()
    
    if(input$run && input$showResults == "original"){
      
      result = datatable(res$resultsAllOriginal$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
    
    
     if(input$run && input$showResults == "transformed"){
      
      result = datatable(res$resultsAllTransformed$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
    
    
     if(input$run && input$showResults == "transformBack"){
      
      result = datatable(res$resultsAllTransformBack$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
    }
result
```

## 4.2. Gender 1

### 4.2.1. Step 1: Coefficient of variation results
```{r}
if(input$subsetAnalysis){
      res = analysisOfVarinceResults()
      
      if(input$run && input$showResults == "original"){
        
        result = datatable(res$resultsOriginalGender1$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      else if(input$run && input$showResults == "transformed"){
        
        
        result = datatable(res$resultsTransformedGender1$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformBack"){
        
        result =  datatable(res$resultsTransformBackGender1$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      result
      
    }

```


### 4.2.2. Step 2: Analysis of variance table
```{r}
  if(input$subsetAnalysis){
      
      res = analysisOfVarinceResults()
      
      if(input$run && input$showResults == "original"){
        
        result =  datatable(res$resultsOriginalGender1$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformed"){
        
        result = datatable(res$resultsTransformedGender1$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformBack"){
        
        result = datatable(res$resultsTransformBackGender1$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      result
    }

```


### 4.2.3. Step 3: Quality measures
```{r}
if(input$subsetAnalysis){
      
      res = analysisOfVarinceResults()
      
      
      if(input$run && input$showResults == "original"){
        
        result = datatable(res$resultsOriginalGender1$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformed"){
        
        result = datatable(res$resultsTransformedGender1$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformBack"){
        
        result = datatable(res$resultsTransformBackGender1$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      result
      
    }

```


## 4.3. Gender 2


### 4.3.1. Step 1: Coefficient of variation results
```{r}
if(input$subsetAnalysis){
      
      result = res = analysisOfVarinceResults()
      
      if(input$run && input$showResults == "original"){
        
        result =  datatable(res$resultsOriginalGender2$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformed"){
        
        result =   datatable(res$resultsTransformedGender2$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformBack"){
        
        result = datatable(res$resultsTransformBackGender2$CVresults, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
     result 
    }

```


### 4.3.2. Step 2: Analysis of variance table
```{r}
  if(input$subsetAnalysis){
      res = analysisOfVarinceResults()
      
      if(input$run && input$showResults == "original"){
        
        result =  datatable(res$resultsOriginalGender2$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformed"){
        
        result = datatable(res$resultsTransformedGender2$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformBack"){
        
        result = datatable(res$resultsTransformBackGender2$CVTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
    result
      }

```


### 4.3.3. Step 3: Quality measures
```{r}

 if(input$subsetAnalysis){
      
      res = analysisOfVarinceResults()
      
      
      if(input$run && input$showResults == "original"){
        
        result =  datatable(res$resultsOriginalGender2$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformed"){
        
        result =  datatable(res$resultsTransformedGender2$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
      
      
      else if(input$run && input$showResults == "transformBack"){
        
        result = datatable(res$resultsTransformBackGender2$errorTable, extensions = c('Buttons','KeyTable', 'Responsive'), options = list(           dom = 'Bfrtip',buttons = list('copy', 'print', list(extend = 'collection',           buttons = c('csv', 'excel', 'pdf'),text = 'Download')), keys = TRUE))
      }
    result
      }
```


